---
title: "small-lm-analysis"
author: "Zachary Houghton"
date: "2025-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brms)
```

# Prompt 1: "This next phrase must be formed compositionally: "

## Prep Data

Read in our model data

```{r}
corpus = read_csv('../Data/nonce_and_attested_binoms.csv')

model_smolm1 = read_csv('../Data/0-smolm-autoreg-bpe-babylm-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  
model_smolm2 = read_csv('../Data/0-smolm-autoreg-bpe-babylm-seed_211-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form + 0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm3 = read_csv('../Data/0-smolm-autoreg-bpe-babylm-seed_1024-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm1 = model_smolm1 %>%
  rename(no_final_stress = `*BStress`)

model_smolm2 = model_smolm2 %>%
  rename(no_final_stress = `*BStress`)

model_smolm3 = model_smolm3 %>%
  rename(no_final_stress = `*BStress`)
```

```{r}

model_smolm1_attested = model_smolm1 %>%
  filter(Attested == 1)

model_smolm1_novel = model_smolm1 %>%
  filter(Attested == 0)

model_smolm2_attested = model_smolm2 %>%
  filter(Attested == 1)

model_smolm2_novel = model_smolm2 %>%
  filter(Attested == 0)

model_smolm3_attested = model_smolm3 %>%
  filter(Attested == 1)

model_smolm3_novel = model_smolm3 %>%
  filter(Attested == 0)

#bigram_counts = read_csv('../Data/olmo_bigram_freqs.csv')
#onegram_corpus_size = 1560674367427
#count_and = 43378012800

#bigram_counts = bigram_counts %>%
  #mutate(ngram = str_replace_all(ngram, '\t', ' '))
# 
# data_main_model = data_main_model %>%
#   mutate(bigram1_alpha = paste0(Word1, ' and'),
#          bigram2_alpha = paste0('and ', Word2),
#          bigram1_nonalpha = paste0(Word2, ' and'),
#          bigram2_nonalpha = paste0('and ', Word1)
#          )
# 
# data_main_model = data_main_model %>%
#   left_join(bigram_counts, by = c('bigram1_alpha' = 'ngram')) %>%
#   rename(count_bigram1_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_alpha' = 'ngram')) %>%
#   rename(count_bigram2_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram1_nonalpha' = 'ngram')) %>%
#   rename(count_bigram1_nonalpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_nonalpha' = 'ngram')) %>%
#   rename(count_bigram2_nonalpha = count) %>%
#   mutate(corpus_size = onegram_corpus_size) %>%
#   mutate(count_and = count_and)

# 
# data_main_model = data_main_model %>%
#   mutate(bigram_prob_alpha = (Word1_freq / corpus_size) * (count_bigram1_alpha / Word1_freq) * (count_bigram2_alpha / count_and),
#          bigram_prob_nonalpha = (Word2_freq / corpus_size) * (count_bigram1_nonalpha / Word2_freq) * (count_bigram2_nonalpha / count_and),
#          log_bigram_prob_alpha = log(bigram_prob_alpha),
#          log_bigram_prob_nonalpha = log(bigram_prob_nonalpha),
#          log_bigram_odds_ratio = log_bigram_prob_alpha - log_bigram_prob_nonalpha)
#
```

## Analyses

### Attested Binomials

```{r}
options(contrasts = c("contr.sum","contr.sum"))
prior_probs = c(
  prior(student_t(3, 0, 1), class = 'Intercept'),
  prior(student_t(3, 0, 1), class = 'sigma'),
  prior(student_t(3, 0, 1), class = 'b')
)
model_smolm1_full_df = model_smolm1 
#model_smolm1_full_df$GenPref = scale(model_smolm1_full_df$GenPref)[,1]
#model_smolm1_full_df$RelFreq = scale(model_smolm1_full_df$RelFreq)[,1]
#model_smolm1_full_df$LogCenteredOverallFreq = scale(log(model_smolm1_full_df$OverallFreq + 0.000000001))[,1]


model_smolm1_full = brm(log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm1_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_full'
                      )

model_smolm1_full_df = model_smolm1
model_smolm1_full_df$LogOverallFreq = scale(log(model_smolm1_full_df$OverallFreq + 1))[,1]

model_smolm1_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm1_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_full_log_overall_freq_scaled'
                      )



model_smolm2_full_df = model_smolm2
model_smolm2_full_df$LogOverallFreq = scale(log(model_smolm2_full_df$OverallFreq + 1))[,1]

model_smolm2_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm2_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2_full_log_overall_freq_scaled'
                      )



model_smolm3_full_df = model_smolm3
model_smolm3_full_df$LogOverallFreq = scale(log(model_smolm3_full_df$OverallFreq + 1))[,1]

model_smolm3_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm3_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3_full_log_overall_freq_scaled'
                      )




fixef(model_smolm1_full_log_overall_freq_scaled)
fixef(model_smolm2_full_log_overall_freq_scaled)
fixef(model_smolm3_full_log_overall_freq_scaled)


model_smolm1_attested$LogOverallFreq = scale(log(model_smolm1_attested$OverallFreq))

model_smolm1_attested_model = brm( log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm1_attested,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_attested_model'
                      )


model_smolm2_attested$LogOverallFreq = scale(log(model_smolm2_attested$OverallFreq))

model_smolm2_attested_model = brm( log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm2_attested,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2_attested_model'
                      )

model_smolm3_attested$LogOverallFreq = scale(log(model_smolm3_attested$OverallFreq))

model_smolm3_attested_model = brm( log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm3_attested,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3_attested_model'
                      )


fixef(model_smolm1_attested_model)
fixef(model_smolm2_attested_model)
fixef(model_smolm3_attested_model)
```

### Novel Binomials

```{r}
options(contrasts = c("contr.sum","contr.sum"))
prior_probs = c(
  prior(student_t(3, 0, 1), class = 'Intercept'),
  prior(student_t(3, 0, 1), class = 'sigma'),
  prior(student_t(3, 0, 1), class = 'b')
)


model_smolm1_novel_model = brm(log_odds ~ GenPref,
                       data = model_smolm1_novel,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm1_novel_model'
                      )

model_smolm2_novel_model = brm(log_odds ~ GenPref,
                       data = model_smolm2_novel,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm2_novel_model'
                      )

model_smolm3_novel_model = brm(log_odds ~ GenPref,
                       data = model_smolm3_novel,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm3_novel_model'
                      )

fixef(model_smolm1_novel_model)
fixef(model_smolm2_novel_model)
fixef(model_smolm3_novel_model)



ce <- conditional_effects(model_smolm1_novel_model, effects = "GenPref",
                          re_formula = NA)   # fixed effects only
p = plot(ce, points = TRUE)
p[[1]] + theme_bw()



model_smolm1_novel_model_test = brm(log_odds ~ scale(GenPref),
                       data = model_smolm1_novel,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm1_novel_model_test2'
                      )

fixef(model_smolm1_novel_model_test)
ce <- conditional_effects(model_smolm1_novel_model_test, effects = "GenPref",
                          re_formula = NA)   # fixed effects only
p = plot(ce, points = TRUE)
p[[1]] + 
  theme_bw() +
  xlab('Abstract Preferences') +
  ylab('Log Odds Alphabetical to Nonalphabetical') +
  theme(axis.title.x = element_text(size = 14),
        axis.title.y = element_text(size = 14)) 

```

# Prompt 2: "Next item: "

## Prep Data

Read in our model data

```{r}

model_smolm1b = read_csv('../Data/1-smolm-autoreg-bpe-babylm-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  
model_smolm2b = read_csv('../Data/1-smolm-autoreg-bpe-babylm-seed_211-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form + 0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm3b = read_csv('../Data/1-smolm-autoreg-bpe-babylm-seed_1024-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm1b = model_smolm1b %>%
  rename(no_final_stress = `*BStress`)

model_smolm2b = model_smolm2b %>%
  rename(no_final_stress = `*BStress`)

model_smolm3b = model_smolm3b %>%
  rename(no_final_stress = `*BStress`)
```

```{r}

model_smolm1_attestedb = model_smolm1b %>%
  filter(Attested == 1)

model_smolm1_novelb = model_smolm1b %>%
  filter(Attested == 0)

model_smolm2_attestedb = model_smolm2b %>%
  filter(Attested == 1)

model_smolm2_novelb = model_smolm2b %>%
  filter(Attested == 0)

model_smolm3_attestedb = model_smolm3b %>%
  filter(Attested == 1)

model_smolm3_novelb = model_smolm3b %>%
  filter(Attested == 0)

#bigram_counts = read_csv('../Data/olmo_bigram_freqs.csv')
#onegram_corpus_size = 1560674367427
#count_and = 43378012800

#bigram_counts = bigram_counts %>%
  #mutate(ngram = str_replace_all(ngram, '\t', ' '))
# 
# data_main_model = data_main_model %>%
#   mutate(bigram1_alpha = paste0(Word1, ' and'),
#          bigram2_alpha = paste0('and ', Word2),
#          bigram1_nonalpha = paste0(Word2, ' and'),
#          bigram2_nonalpha = paste0('and ', Word1)
#          )
# 
# data_main_model = data_main_model %>%
#   left_join(bigram_counts, by = c('bigram1_alpha' = 'ngram')) %>%
#   rename(count_bigram1_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_alpha' = 'ngram')) %>%
#   rename(count_bigram2_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram1_nonalpha' = 'ngram')) %>%
#   rename(count_bigram1_nonalpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_nonalpha' = 'ngram')) %>%
#   rename(count_bigram2_nonalpha = count) %>%
#   mutate(corpus_size = onegram_corpus_size) %>%
#   mutate(count_and = count_and)

# 
# data_main_model = data_main_model %>%
#   mutate(bigram_prob_alpha = (Word1_freq / corpus_size) * (count_bigram1_alpha / Word1_freq) * (count_bigram2_alpha / count_and),
#          bigram_prob_nonalpha = (Word2_freq / corpus_size) * (count_bigram1_nonalpha / Word2_freq) * (count_bigram2_nonalpha / count_and),
#          log_bigram_prob_alpha = log(bigram_prob_alpha),
#          log_bigram_prob_nonalpha = log(bigram_prob_nonalpha),
#          log_bigram_odds_ratio = log_bigram_prob_alpha - log_bigram_prob_nonalpha)
#
```

## Analyses

### Attested Binomials

```{r}


model_smolm1b_full_df = model_smolm1b
model_smolm1b_full_df$LogOverallFreq = scale(log(model_smolm1b_full_df$OverallFreq + 1))[,1]

model_smolm1b_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm1b_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1b_full_log_overall_freq_scaled'
                      )
fixef(model_smolm1b_full_log_overall_freq_scaled)

model_smolm2b_full_df = model_smolm2b
model_smolm2b_full_df$LogOverallFreq = scale(log(model_smolm2b_full_df$OverallFreq + 1))[,1]

model_smolm2b_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm2b_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2b_full_log_overall_freq_scaled'
                      )
fixef(model_smolm2b_full_log_overall_freq_scaled)

model_smolm3b_full_df = model_smolm3b
model_smolm3b_full_df$LogOverallFreq = scale(log(model_smolm3b_full_df$OverallFreq + 1))[,1]

model_smolm3b_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm3b_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3b_full_log_overall_freq_scaled'
                      )
fixef(model_smolm3b_full_log_overall_freq_scaled)

model_smolm1_attested_modelb = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm1_attestedb,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_attested_modelb'
                      )

model_smolm2_attested_modelb = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm2_attestedb,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2_attested_modelb'
                      )


model_smolm3_attested_modelb = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm3_attestedb,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3_attested_modelb'
                      )


fixef(model_smolm1_attested_modelb)
fixef(model_smolm2_attested_modelb)
fixef(model_smolm3_attested_modelb)
```

### Novel Binomials

```{r}


model_smolm1_novel_modelb = brm(log_odds ~ GenPref,
                       data = model_smolm1_novelb,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm1_novel_modelb'
                      )

model_smolm2_novel_modelb = brm(log_odds ~ GenPref,
                       data = model_smolm2_novelb,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm2_novel_modelb'
                      )

model_smolm3_novel_modelb = brm(log_odds ~ GenPref,
                       data = model_smolm3_novelb,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm3_novel_modelb'
                      )

fixef(model_smolm1_novel_modelb)
fixef(model_smolm2_novel_modelb)
fixef(model_smolm3_novel_modelb)
```


# Prompt 3: "example: "

## Prep Data

Read in our model data

```{r}

model_smolm1c = read_csv('../Data/2-smolm-autoreg-bpe-babylm-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  
model_smolm2c = read_csv('../Data/2-smolm-autoreg-bpe-babylm-seed_211-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form + 0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm3c = read_csv('../Data/2-smolm-autoreg-bpe-babylm-seed_1024-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm1c = model_smolm1c %>%
  rename(no_final_stress = `*BStress`)

model_smolm2c = model_smolm2c %>%
  rename(no_final_stress = `*BStress`)

model_smolm3c = model_smolm3c %>%
  rename(no_final_stress = `*BStress`)
```

```{r}

model_smolm1_attestedc = model_smolm1c %>%
  filter(Attested == 1)

model_smolm1_novelc = model_smolm1c %>%
  filter(Attested == 0)

model_smolm2_attestedc = model_smolm2c %>%
  filter(Attested == 1)

model_smolm2_novelc = model_smolm2c %>%
  filter(Attested == 0)

model_smolm3_attestedc = model_smolm3c %>%
  filter(Attested == 1)

model_smolm3_novelc = model_smolm3c %>%
  filter(Attested == 0)

#bigram_counts = read_csv('../Data/olmo_bigram_freqs.csv')
#onegram_corpus_size = 1560674367427
#count_and = 43378012800

#bigram_counts = bigram_counts %>%
  #mutate(ngram = str_replace_all(ngram, '\t', ' '))
# 
# data_main_model = data_main_model %>%
#   mutate(bigram1_alpha = paste0(Word1, ' and'),
#          bigram2_alpha = paste0('and ', Word2),
#          bigram1_nonalpha = paste0(Word2, ' and'),
#          bigram2_nonalpha = paste0('and ', Word1)
#          )
# 
# data_main_model = data_main_model %>%
#   left_join(bigram_counts, by = c('bigram1_alpha' = 'ngram')) %>%
#   rename(count_bigram1_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_alpha' = 'ngram')) %>%
#   rename(count_bigram2_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram1_nonalpha' = 'ngram')) %>%
#   rename(count_bigram1_nonalpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_nonalpha' = 'ngram')) %>%
#   rename(count_bigram2_nonalpha = count) %>%
#   mutate(corpus_size = onegram_corpus_size) %>%
#   mutate(count_and = count_and)

# 
# data_main_model = data_main_model %>%
#   mutate(bigram_prob_alpha = (Word1_freq / corpus_size) * (count_bigram1_alpha / Word1_freq) * (count_bigram2_alpha / count_and),
#          bigram_prob_nonalpha = (Word2_freq / corpus_size) * (count_bigram1_nonalpha / Word2_freq) * (count_bigram2_nonalpha / count_and),
#          log_bigram_prob_alpha = log(bigram_prob_alpha),
#          log_bigram_prob_nonalpha = log(bigram_prob_nonalpha),
#          log_bigram_odds_ratio = log_bigram_prob_alpha - log_bigram_prob_nonalpha)
#
```

## Analyses

### Attested Binomials

```{r}

model_smolm1c_full_df = model_smolm1c
model_smolm1c_full_df$LogOverallFreq = scale(log(model_smolm1c_full_df$OverallFreq + 1))[,1]

model_smolm1c_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm1c_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1c_full_log_overall_freq_scaled'
                      )
fixef(model_smolm1c_full_log_overall_freq_scaled)

model_smolm2c_full_df = model_smolm2c
model_smolm2c_full_df$LogOverallFreq = scale(log(model_smolm2c_full_df$OverallFreq + 1))[,1]

model_smolm2c_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm2c_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2c_full_log_overall_freq_scaled'
                      )
fixef(model_smolm2c_full_log_overall_freq_scaled)

model_smolm3c_full_df = model_smolm3c
model_smolm3c_full_df$LogOverallFreq = scale(log(model_smolm3c_full_df$OverallFreq + 1))[,1]

model_smolm3c_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm3c_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3c_full_log_overall_freq_scaled'
                      )
fixef(model_smolm3c_full_log_overall_freq_scaled)

model_smolm1_attested_modelc = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm1_attestedc,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_attested_modelc'
                      )

model_smolm2_attested_modelc = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm2_attestedc,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2_attested_modelc'
                      )


model_smolm3_attested_modelc = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm3_attestedc,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3_attested_modelc'
                      )


fixef(model_smolm1_attested_modelc)
fixef(model_smolm2_attested_modelc)
fixef(model_smolm3_attested_modelc)
```

### Novel Binomials

```{r}


model_smolm1_novel_modelc = brm(log_odds ~ GenPref,
                       data = model_smolm1_novelc,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm1_novel_modelc'
                      )

model_smolm2_novel_modelc = brm(log_odds ~ GenPref,
                       data = model_smolm2_novelc,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm2_novel_modelc'
                      )

model_smolm3_novel_modelc = brm(log_odds ~ GenPref,
                       data = model_smolm3_novelc,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm3_novel_modelc'
                      )

fixef(model_smolm1_novel_modelc)
fixef(model_smolm2_novel_modelc)
fixef(model_smolm3_novel_modelc)
```



# Prompt 4: "instance: "

## Prep Data

Read in our model data

```{r}

model_smolm1d = read_csv('../Data/3-smolm-autoreg-bpe-babylm-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  
model_smolm2d = read_csv('../Data/3-smolm-autoreg-bpe-babylm-seed_211-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form + 0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm3d = read_csv('../Data/3-smolm-autoreg-bpe-babylm-seed_1024-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm1d = model_smolm1d %>%
  rename(no_final_stress = `*BStress`)

model_smolm2d = model_smolm2d %>%
  rename(no_final_stress = `*BStress`)

model_smolm3d = model_smolm3d %>%
  rename(no_final_stress = `*BStress`)
```

```{r}

model_smolm1_attestedd = model_smolm1d %>%
  filter(Attested == 1)

model_smolm1_noveld = model_smolm1d %>%
  filter(Attested == 0)

model_smolm2_attestedd = model_smolm2d %>%
  filter(Attested == 1)

model_smolm2_noveld = model_smolm2d %>%
  filter(Attested == 0)

model_smolm3_attestedd = model_smolm3d %>%
  filter(Attested == 1)

model_smolm3_noveld = model_smolm3d %>%
  filter(Attested == 0)

#bigram_counts = read_csv('../Data/olmo_bigram_freqs.csv')
#onegram_corpus_size = 1560674367427
#count_and = 43378012800

#bigram_counts = bigram_counts %>%
  #mutate(ngram = str_replace_all(ngram, '\t', ' '))
# 
# data_main_model = data_main_model %>%
#   mutate(bigram1_alpha = paste0(Word1, ' and'),
#          bigram2_alpha = paste0('and ', Word2),
#          bigram1_nonalpha = paste0(Word2, ' and'),
#          bigram2_nonalpha = paste0('and ', Word1)
#          )
# 
# data_main_model = data_main_model %>%
#   left_join(bigram_counts, by = c('bigram1_alpha' = 'ngram')) %>%
#   rename(count_bigram1_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_alpha' = 'ngram')) %>%
#   rename(count_bigram2_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram1_nonalpha' = 'ngram')) %>%
#   rename(count_bigram1_nonalpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_nonalpha' = 'ngram')) %>%
#   rename(count_bigram2_nonalpha = count) %>%
#   mutate(corpus_size = onegram_corpus_size) %>%
#   mutate(count_and = count_and)

# 
# data_main_model = data_main_model %>%
#   mutate(bigram_prob_alpha = (Word1_freq / corpus_size) * (count_bigram1_alpha / Word1_freq) * (count_bigram2_alpha / count_and),
#          bigram_prob_nonalpha = (Word2_freq / corpus_size) * (count_bigram1_nonalpha / Word2_freq) * (count_bigram2_nonalpha / count_and),
#          log_bigram_prob_alpha = log(bigram_prob_alpha),
#          log_bigram_prob_nonalpha = log(bigram_prob_nonalpha),
#          log_bigram_odds_ratio = log_bigram_prob_alpha - log_bigram_prob_nonalpha)
#
```

## Analyses

### Attested Binomials

```{r}

model_smolm1d_full_df = model_smolm1d
model_smolm1d_full_df$LogOverallFreq = scale(log(model_smolm1d_full_df$OverallFreq + 1))[,1]

model_smolm1d_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm1d_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1d_full_log_overall_freq_scaled'
                      )

model_smolm2d_full_df = model_smolm2d
model_smolm2d_full_df$LogOverallFreq = scale(log(model_smolm2d_full_df$OverallFreq + 1))[,1]

model_smolm2d_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm2d_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2d_full_log_overall_freq_scaled'
                      )



model_smolm3d_full_df = model_smolm3d
model_smolm3d_full_df$LogOverallFreq = scale(log(model_smolm3d_full_df$OverallFreq + 1))[,1]

model_smolm3d_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm3d_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3d_full_log_overall_freq_scaled'
                      )

fixef(model_smolm3d_full_log_overall_freq_scaled)


model_smolm1_attested_modeld = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm1_attestedd,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_attested_modeld'
                      )

model_smolm2_attested_modeld = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm2_attestedd,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2_attested_modeld'
                      )


model_smolm3_attested_modeld = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm3_attestedd,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3_attested_modeld'
                      )


fixef(model_smolm1_attested_modeld)
fixef(model_smolm2_attested_modeld)
fixef(model_smolm3_attested_modeld)
```

### Novel Binomials

```{r}


model_smolm1_novel_modeld = brm(log_odds ~ GenPref,
                       data = model_smolm1_noveld,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm1_novel_modeld'
                      )

model_smolm2_novel_modeld = brm(log_odds ~ GenPref,
                       data = model_smolm2_noveld,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm2_novel_modeld'
                      )

model_smolm3_novel_modeld = brm(log_odds ~ GenPref,
                       data = model_smolm3_noveld,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm3_novel_modeld'
                      )

fixef(model_smolm1_novel_modeld)
fixef(model_smolm2_novel_modeld)
fixef(model_smolm3_novel_modeld)
```



# Prompt 5: "try this: "

## Prep Data

Read in our model data

```{r}

model_smolm1e = read_csv('../Data/4-smolm-autoreg-bpe-babylm-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  
model_smolm2e = read_csv('../Data/4-smolm-autoreg-bpe-babylm-seed_211-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form + 0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm3e = read_csv('../Data/4-smolm-autoreg-bpe-babylm-seed_1024-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm1e = model_smolm1e %>%
  rename(no_final_stress = `*BStress`)

model_smolm2e = model_smolm2e %>%
  rename(no_final_stress = `*BStress`)

model_smolm3e = model_smolm3e %>%
  rename(no_final_stress = `*BStress`)
```

```{r}

model_smolm1_attestede = model_smolm1e %>%
  filter(Attested == 1)

model_smolm1_novele = model_smolm1e %>%
  filter(Attested == 0)

model_smolm2_attestede = model_smolm2e %>%
  filter(Attested == 1)

model_smolm2_novele = model_smolm2e %>%
  filter(Attested == 0)

model_smolm3_attestede = model_smolm3e %>%
  filter(Attested == 1)

model_smolm3_novele = model_smolm3e %>%
  filter(Attested == 0)

#bigram_counts = read_csv('../Data/olmo_bigram_freqs.csv')
#onegram_corpus_size = 1560674367427
#count_and = 43378012800

#bigram_counts = bigram_counts %>%
  #mutate(ngram = str_replace_all(ngram, '\t', ' '))
# 
# data_main_model = data_main_model %>%
#   mutate(bigram1_alpha = paste0(Word1, ' and'),
#          bigram2_alpha = paste0('and ', Word2),
#          bigram1_nonalpha = paste0(Word2, ' and'),
#          bigram2_nonalpha = paste0('and ', Word1)
#          )
# 
# data_main_model = data_main_model %>%
#   left_join(bigram_counts, by = c('bigram1_alpha' = 'ngram')) %>%
#   rename(count_bigram1_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_alpha' = 'ngram')) %>%
#   rename(count_bigram2_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram1_nonalpha' = 'ngram')) %>%
#   rename(count_bigram1_nonalpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_nonalpha' = 'ngram')) %>%
#   rename(count_bigram2_nonalpha = count) %>%
#   mutate(corpus_size = onegram_corpus_size) %>%
#   mutate(count_and = count_and)

# 
# data_main_model = data_main_model %>%
#   mutate(bigram_prob_alpha = (Word1_freq / corpus_size) * (count_bigram1_alpha / Word1_freq) * (count_bigram2_alpha / count_and),
#          bigram_prob_nonalpha = (Word2_freq / corpus_size) * (count_bigram1_nonalpha / Word2_freq) * (count_bigram2_nonalpha / count_and),
#          log_bigram_prob_alpha = log(bigram_prob_alpha),
#          log_bigram_prob_nonalpha = log(bigram_prob_nonalpha),
#          log_bigram_odds_ratio = log_bigram_prob_alpha - log_bigram_prob_nonalpha)
#
```

## Analyses

### Attested Binomials

```{r}
model_smolm1e_full_df = model_smolm1e
model_smolm1e_full_df$LogOverallFreq = scale(log(model_smolm1e_full_df$OverallFreq + 1))[,1]

model_smolm1e_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm1e_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1e_full_log_overall_freq_scaled'
                      )
fixef(model_smolm1e_full_log_overall_freq_scaled)

model_smolm2e_full_df = model_smolm2e
model_smolm2e_full_df$LogOverallFreq = scale(log(model_smolm2e_full_df$OverallFreq + 1))[,1]

model_smolm2e_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm2e_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2e_full_log_overall_freq_scaled'
                      )
fixef(model_smolm2e_full_log_overall_freq_scaled)


model_smolm3e_full_df = model_smolm3e
model_smolm3e_full_df$LogOverallFreq = scale(log(model_smolm3e_full_df$OverallFreq + 1))[,1]

model_smolm3e_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm3e_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3e_full_log_overall_freq_scaled'
                      )
fixef(model_smolm3e_full_log_overall_freq_scaled)

model_smolm1_attested_modele = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm1_attestede,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_attested_modele'
                      )

model_smolm2_attested_modele = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm2_attestede,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2_attested_modele'
                      )


model_smolm3_attested_modele = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm3_attestede,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3_attested_modele'
                      )


fixef(model_smolm1_attested_modele)
fixef(model_smolm2_attested_modele)
fixef(model_smolm3_attested_modele)
```

### Novel Binomials

```{r}


model_smolm1_novel_modele = brm(log_odds ~ GenPref,
                       data = model_smolm1_novele,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm1_novel_modele'
                      )

model_smolm2_novel_modele = brm(log_odds ~ GenPref,
                       data = model_smolm2_novele,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm2_novel_modele'
                      )

model_smolm3_novel_modele = brm(log_odds ~ GenPref,
                       data = model_smolm3_novele,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm3_novel_modele'
                      )

fixef(model_smolm1_novel_modele)
fixef(model_smolm2_novel_modele)
fixef(model_smolm3_novel_modele)
```



# Prompt 6: "[BOS]"

## Prep Data

Read in our model data

```{r}

model_smolm1f = read_csv('../Data/5-smolm-autoreg-bpe-babylm-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  
model_smolm2f = read_csv('../Data/5-smolm-autoreg-bpe-babylm-seed_211-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form + 0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm3f = read_csv('../Data/5-smolm-autoreg-bpe-babylm-seed_1024-1e-3_binom_ordering_prefs.csv') %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = case_when(Attested == 1 ~ RelFreq - 0.5,
                             Attested == 0 ~ 0)) %>%
  mutate(ProbAandB = exp(`Alpha Probs`) / (exp(`Alpha Probs`) + exp(`Nonalpha Probs`))) %>%
  mutate(log_odds = `Alpha Probs` - `Nonalpha Probs`)
  

model_smolm1f = model_smolm1f %>%
  rename(no_final_stress = `*BStress`)

model_smolm2f = model_smolm2f %>%
  rename(no_final_stress = `*BStress`)

model_smolm3f = model_smolm3f %>%
  rename(no_final_stress = `*BStress`)
```

```{r}

model_smolm1_attestedf = model_smolm1f %>%
  filter(Attested == 1)

model_smolm1_novelf = model_smolm1f %>%
  filter(Attested == 0)

model_smolm2_attestedf = model_smolm2f %>%
  filter(Attested == 1)

model_smolm2_novelf = model_smolm2f %>%
  filter(Attested == 0)

model_smolm3_attestedf = model_smolm3f %>%
  filter(Attested == 1)

model_smolm3_novelf = model_smolm3f %>%
  filter(Attested == 0)

#bigram_counts = read_csv('../Data/olmo_bigram_freqs.csv')
#onegram_corpus_size = 1560674367427
#count_and = 43378012800

#bigram_counts = bigram_counts %>%
  #mutate(ngram = str_replace_all(ngram, '\t', ' '))
# 
# data_main_model = data_main_model %>%
#   mutate(bigram1_alpha = paste0(Word1, ' and'),
#          bigram2_alpha = paste0('and ', Word2),
#          bigram1_nonalpha = paste0(Word2, ' and'),
#          bigram2_nonalpha = paste0('and ', Word1)
#          )
# 
# data_main_model = data_main_model %>%
#   left_join(bigram_counts, by = c('bigram1_alpha' = 'ngram')) %>%
#   rename(count_bigram1_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_alpha' = 'ngram')) %>%
#   rename(count_bigram2_alpha = count) %>%
#   left_join(bigram_counts, by = c('bigram1_nonalpha' = 'ngram')) %>%
#   rename(count_bigram1_nonalpha = count) %>%
#   left_join(bigram_counts, by = c('bigram2_nonalpha' = 'ngram')) %>%
#   rename(count_bigram2_nonalpha = count) %>%
#   mutate(corpus_size = onegram_corpus_size) %>%
#   mutate(count_and = count_and)

# 
# data_main_model = data_main_model %>%
#   mutate(bigram_prob_alpha = (Word1_freq / corpus_size) * (count_bigram1_alpha / Word1_freq) * (count_bigram2_alpha / count_and),
#          bigram_prob_nonalpha = (Word2_freq / corpus_size) * (count_bigram1_nonalpha / Word2_freq) * (count_bigram2_nonalpha / count_and),
#          log_bigram_prob_alpha = log(bigram_prob_alpha),
#          log_bigram_prob_nonalpha = log(bigram_prob_nonalpha),
#          log_bigram_odds_ratio = log_bigram_prob_alpha - log_bigram_prob_nonalpha)
#
```

## Analyses

### Attested Binomials

```{r}
model_smolm1f_full_df = model_smolm1f
model_smolm1f_full_df$LogOverallFreq = scale(log(model_smolm1f_full_df$OverallFreq + 1))[,1]

model_smolm1f_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm1f_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1f_full_log_overall_freq_scaled'
                      )

model_smolm2f_full_df = model_smolm2f
model_smolm2f_full_df$LogOverallFreq = scale(log(model_smolm2f_full_df$OverallFreq + 1))[,1]

model_smolm2f_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm2f_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2f_full_log_overall_freq_scaled'
                      )

model_smolm3f_full_df = model_smolm3f
model_smolm3f_full_df$LogOverallFreq = scale(log(model_smolm3f_full_df$OverallFreq + 1))[,1]

model_smolm3f_full_log_overall_freq_scaled = brm(log_odds ~ GenPref + RelFreq + GenPref:LogOverallFreq + RelFreq:LogOverallFreq + LogOverallFreq,
                       data = model_smolm3f_full_df,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3f_full_log_overall_freq_scaled'
                      )

fixef(model_smolm1f_full_log_overall_freq_scaled)
model_smolm1_attested_modelf = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm1_attestedf,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm1_attested_modelf'
                      )

model_smolm2_attested_modelf = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm2_attestedf,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm2_attested_modelf'
                      )


model_smolm3_attested_modelf = brm( log_odds ~ GenPref + RelFreq + GenPref:OverallFreq + RelFreq:OverallFreq + OverallFreq,
                       data = model_smolm3_attestedf,
                       prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       control = list(adapt_delta=0.99, max_treedepth = 20),
                       file = '../Data/model_smolm3_attested_modelf'
                      )


fixef(model_smolm1_attested_modelf)
fixef(model_smolm2_attested_modelf)
fixef(model_smolm3_attested_modelf)
```

### Novel Binomials

```{r}


model_smolm1_novel_modelf = brm(log_odds ~ GenPref,
                       data = model_smolm1_novelf,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm1_novel_modelf'
                      )

model_smolm2_novel_modelf = brm(log_odds ~ GenPref,
                       data = model_smolm2_novelf,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm2_novel_modelf'
                      )

model_smolm3_novel_modelf = brm(log_odds ~ GenPref,
                       data = model_smolm3_novelf,
                       prior = prior_probs,
                       iter = 6000,
                       warmup = 3000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_smolm3_novel_modelf'
                      )

fixef(model_smolm1_novel_modelf)
fixef(model_smolm2_novel_modelf)
fixef(model_smolm3_novel_modelf)
```


# Qing's model, all prompts

```{r}
corpus = read_csv('../Data/nonce_and_attested_binoms.csv')

small_attested = read_csv('../Data/SMALL_MODEL_BINOMIAL_PREFERENCES.csv') %>%
  mutate(log_odds = preference) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(RelFreq = RelFreq - 0.5) %>%
  mutate(OverallFreq = log(OverallFreq)) %>%
  filter(Attested == 1) 



small_novel = read_csv('../Data/SMALL_MODEL_BINOMIAL_PREFERENCES.csv') %>%
  mutate(log_odds = preference) %>%
  separate(binom, c('Word1', 'and', 'Word2'), remove = F, sep = ' ') %>%
  select(-and) %>%
  #mutate(across(2:3, tolower)) %>%
  left_join(corpus) %>%
  mutate(checkpoint = 'main') %>%
  mutate(y_vals = 0.02191943 + 0.23925834*Form +  0.24889543*Percept +  0.41836997*Culture +   0.25967334*Power +  0.01867604*Intense +  1.30365980*Icon +   0.08553552*Freq +  0.15241566*Len - 0.19381657*Lapse +  0.36019221*`*BStress`) %>%
  mutate(GenPref = 1/(1+exp(-1*y_vals))) %>%
  mutate(GenPref = GenPref - 0.5) %>%
  mutate(OverallFreq = log(OverallFreq)) %>%
  filter(Attested == 0)

small_attested$prompt = factor(small_attested$prompt)
small_novel$prompt = factor(small_novel$prompt)

small_novel = small_novel %>%
  rename(no_final_stress = "*BStress")

small_attested = small_attested %>%
  rename(no_final_stress = "*BStress")

model_small_attested_genpref = brm(log_odds ~ GenPref + RelFreq + OverallFreq + GenPref:OverallFreq + RelFreq:OverallFreq + (GenPref | prompt) + (1 | Alpha),
                       data = small_attested,
                       #prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_small_attested_genpref'
                      )


model_small_novel_genpref = brm(log_odds ~ GenPref + (GenPref | prompt) + (1 | Alpha),
                       data = small_novel,
                       #prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_small_novel_genpref'
                      )


### Vif then backward model selection

model_small_novel_individual_constraints = brm(log_odds ~ Culture + Power + Freq + Len + no_final_stress + Intense + Percept + Lapse + (Culture + Power + Freq + Len + no_final_stress + Intense + Percept + Lapse | prompt) + ( 1 | Alpha),
                       data = small_novel,
                       #prior = prior_probs,
                       iter = 10000,
                       warmup = 5000,
                       chains = 4,
                       cores = 4,
                       #control = list(adapt_delta=0.99, max_treedepth = 15),
                       #control = list(max_treedepth = 20),
                       file = '../Data/model_small_novel_individual_constraints'
                      )
 

fixef(model_small_attested_genpref)
fixef(model_small_novel_genpref)
fixef(model_small_novel_individual_constraints)
car::vif(model_small_attested_individual_constraints)


## Now we do backward model selection? For this project we might actually want to *not* do backward model selection, because constraints that are centered around zero may change after finetuning.


```

